---
title: "Translate your text while you are sleeping!"
author: "Vladimir Zhbanko"
output: html_notebook
---

# Automatic Translations with R and Google Translator API

This document is to create a way to easier read and translate captions in `*.vtt` format

## Section 1 Why do you need it?

### Lecture 1 Introduction

* many files to translate
* automate repetitive tasks
* good quality initial text

## Section 2 Make It!

### Lecture 2 Install R

Re-use lecture about R installation

### Lecture 3 Google Translator API key

* Subscribe to Google Cloud Platform
* Generate your API Key
* Copy your key


















































### Lecture 4 Encrypt your key in R!

It is highly recommended that you encrypt your key! This lecture will show how. Remember that you can learn much more about using Cryptography from my other course **Cryptography is more fun with R!** [course coupon](https://www.udemy.com/keep-your-secrets-under-control/?couponCode=KEEP-SECRET-20)

#### Using Public Key Cryptography to securely store your API Key

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=TRUE}
# Generate your private key and write it to the folder, we assume you will save it to the folder C:/Users/UserName/.ssh/ mac users can adapt the path...
# if necessary install package
# install.packages("openssl"); install.packages("tidyverse")
# loads library open ssl and tidyverse
library(openssl)
library(tidyverse)

# generate private key
rsa_keygen(bits = 5555) %>% write_pem(path = "C:/Users/fxtrams/.ssh/id_api")
# extract and write your public key
read_key(file = "C:/Users/fxtrams/.ssh/id_api", password = "") %>% `[[`("pubkey") %>% write_pem("C:/Users/fxtrams/.ssh/id_api.pub")

```






























Now you have your personal public key which we will use to encrypt the credentials

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
# encrypt your key (I am showing my API key because I will delete it anyhow after creating the course)
## Encrypt with PRIVATE key (e.g. use this code yourself)
"your API key" %>% 
  # serialize the object
  serialize(connection = NULL) %>% 
  # encrypt the object
  encrypt_envelope("C:/Users/fxtrams/.ssh/id_api.pub") %>% 
  # write encrypted data to File
  write_rds("api_key.enc.rds")

```

Now we have our encrypted key inside our project folder! 

If you will use this script later -> delete API key from your script and feel free to use Version Control Repository!

In the next lecture we will see how to read the key back!

**NOTE:** if you plan to collaborate and use your key by multiple persons with version control check out R package 'secret'. Remember that you can learn how to use it in my course about Cryptography in R!

































### Lecture 5 Translate Hello World!

In this lecture we will do a very simple task! We will see how to use package in R called **translateR**

```{r}
library(openssl)
library(tidyverse)
# to install package in R
install.packages("translate")
library(translate)
citation("translate")


```


























```{r}
# help on the package
?translate

# we need our API key to translate 
out <- read_rds("api_key.enc.rds")

# decrypting the password using public data list and private key
api_key <- decrypt_envelope(out$data, out$iv, out$session, "C:/Users/fxtrams/.ssh/id_api", password = "") %>% 
  unserialize()

# usage: translate(query, source, target, key = get.key())
translate("Mi piace questo corso", "it", "en", key = api_key)

```














## Lecture 8 For loop!

Hello! This lecture I am recording during real usage of my code! I will be translating 23 files in 8 different languages!!!



```{r}
# translate 23 files in .vtt format
source("translateVTT.R")
library(openssl)
library(tidyverse)

# make a list of files to translate
filesToTranslate <-list.files("C:/Users/fxtrams/Downloads/", pattern="*.vtt", full.names=TRUE)
# make a list of languages
languages <- c("fr", "tr", "it", "id", "pt", "es", "ms", "de")
# get api key
out <- read_rds("api_key.enc.rds")

# decrypting the password using public data list and private key
api_key <- decrypt_envelope(out$data, out$iv, out$session, "C:/Users/fxtrams/.ssh/id_api", password = "") %>% unserialize()

# starting time of my job
start_time <- Sys.time()
balance_before <- 275.03

# for loop
for (FILE in filesToTranslate) {
  # for loop for languages
  for (LANG in languages) {
    # translation
    translateVTT(fileName = FILE, sourceLang = "en", destLang = LANG, apikey = api_key)
  }
  
}


# starting time of my job
end_time <- Sys.time()

# how much time did it take?
end_time - start_time
# balance
balance_after <- 256.66

# total cost
cost <- balance_before-balance_after

cost

# result:


```
# conclusion!
1.26 hours and 18.37$ to translate 23 files to 8 different languages...



























```{r}

library(stringr)
library(tidyverse)
library(translateR)

# read file -> it will be a dataframe
t <- read.delim("C:/Users/fxtrams/Downloads/L3.vtt", stringsAsFactors = F)

# extract logical vector indicating which rows containing timestamps
x <- t %>% 
  # detect rows with date time (those for not translate)
  apply(MARGIN = 1, str_detect, pattern = "-->")

# extract only rows containing text (e.g. not containing timestamps) 
txt <- subset.data.frame(t, !x) 
# extract only time stamps
tst <- subset.data.frame(t,  x)

# # write to file for translation (manually)
# txt %>% write.table("translate.txt", row.names = F)

# # write lines for translations
# lns <- txt %>% as.matrix() %>% c()
```

## translate this file using translate API paid service in Google

```{r}
# translate object txt or file in R
# Google, translate column in dataset
google.dataset.out <- translate(dataset = txt,
                                content.field = 'WEBVTT',
                                google.api.key = "api key do not check me in to Version Control!",
                                source.lang = 'en',
                                target.lang = 'es')

# extract only new column
trsltd <- google.dataset.out %>% select(translatedContent)

# give original name
colnames(trsltd) <- "WEBVTT"

# 
```

## place it back...

```{r}
## read this file back
#tsltd <- read.table("translate.txt", encoding = "UTF-8", header = T, stringsAsFactors = F)

# add original row names to the table tsltd
# row.names(tsltd) <- as.numeric(row.names(txt))

# bind rows with original timestamps
abc <- rbind(tst, trsltd)

# order this file back again
bcd <-  abc[ order(as.numeric(row.names(abc))), ] %>% as.character() %>% as.data.frame()

# return original name
colnames(bcd) <- "WEBVTT"

bcd <- as.tibble(bcd)
# add one row
bcd2 <- add_row(bcd, WEBVTT  = "", .before = 1)

# write this file back :_)
write.table(bcd2, "translated.vtt",quote = F, row.names = F)


```


# billing
https://console.cloud.google.com/billing/01B83E-7B64EC-1DD123

```{r}
# translate all courses
source("translateVTT.R")
library(secret)
library(openssl)
library(tidyverse)

# make a list of files to translate
filesToTranslate <-list.files("C:/Users/fxtrams/Downloads/", pattern="*.vtt", full.names=TRUE)
# make a list of languages
languages <- c("fr", "tr", "it", "id", "pt", "es", "ms", "de")
# get api key
api_key <- get_secret("api_google_Translate", key = "C:/Users/fxtrams/.ssh/id_api", vault = "COMMON")

# for loop that translate all languages!
for (fl in filesToTranslate) {
  # another for loop for languages
  for (la in languages) {
    translateVTT(fl, "en", la, api_key)
  }
  
}


```


